# Basic stats

```{r get pipeline version}
pipeSMK <- readLines(file.path(pipelineFolder,"GBS-pipeline.smk"))
pipeVersion <- gsub("##### Version: ","",pipeSMK[grep("##### Version:", pipeSMK)])
```

```{r import barcodesID, results="asis"}
#barcodes <- read.table(file.path(projFolder, "barcodesID.txt"))
#rawsamples <- read.table(file.path(projFolder, "rawsamples"))
#sampleInfo <- read.table(file.path(projFolder, "sampleInfo.txt"), header=TRUE)
pipelineConfig <- readLines(pipelineConfig.file)

sampleSheet.file <- getConfigField("samplesheet-file")
sampleSheet <- read.table(sampleSheet.file, sep="\t", header=TRUE)

sampleInfo.file <- getConfigField("sampleinfo-file")
sampleInfo <- NULL

if(sampleInfo.file!="") sampleInfo <- read.table(sampleInfo.file, sep="\t", header=TRUE)

#projectFolder.info <- trimws(gsub("project-folder:","", pipelineConfig[grep("project-folder:", pipelineConfig)]))
#pipelineFolder.info <- trimws(gsub("pipeline-folder:","", pipelineConfig[grep("pipeline-folder:", pipelineConfig)]))
#pipelineConfig.info <- trimws(gsub("pipeline-config:","", pipelineConfig[grep("pipeline-config:", pipelineConfig)]))

projectFolder.info <- getConfigField("project-folder")
pipelineFolder.info <- getConfigField("pipeline-folder")
pipelineConfig.info <- pipelineConfig.file

groupingNames <- "No file with sample information provided ('sampleinfo-file'-field)!"
numberGroupingNames <- 0
if(!is.null(sampleInfo)){
  groupingNames <- colnames(sampleInfo[,-1])
  numberGroupingNames <- length(groupingNames)
} 

mockSamples <- unique(sampleSheet$sample_name[sampleSheet$useForMock=="YES"])

out <- data.frame(c("Number of raw-samples",
                    "Number of samples (after concatenating)",
                    "No. of grouping information",
                    "Grouping names",
                    "No. of mock reference samples",
                    "Samples used for mock reference",
                    "Used reference genome",
                    "Pipeline version",
                    "Project folder",
                    "Pipeline folder",
                    "Pipeline config",
                    "Enzyme 1 sequence",
                    "Enzyme 2 sequence"),
                  c(length(unique(sampleSheet$rawsample)),
                    length(unique(sampleSheet$sample_name)),
                    numberGroupingNames,
                    groupingNames,
                    length(mockSamples),
                    paste(mockSamples,collapse=", "),
                    basename(refGenome.file),
                    pipeVersion,
                    projectFolder.info,
                    pipelineFolder.info,
                    pipelineConfig.info,
                    enz1,
                    enz2))

out_html <- knitr::kable(out, col.names = NULL, "html")
kable_styling(out_html, "striped", position = "left")
```

## Concatenating stats
As the concatenating is a crucial step in this pipeline, it is important to make this correct. As a double check that things went right, we check how
many raw samples were concatenated per sample. We do this visually. Basically, the height of the bars should correspond to your number of used lanes,
normally 1,2 or 4. Sometimes, you might want to concatenate also more files together (e.g. technical replicates), in this case the bar can be also higher.

```{r import the conc report files}
conc.reports.files <- list.files(file.path(projFolder, "FASTQ", "CONCATENATED"), pattern="*.report")

conc.reports <- list()
conc.reports[[1]] <- read.table(file.path(projFolder, "FASTQ", "CONCATENATED", conc.reports.files[1]))

for(i in 2:length(conc.reports.files)){
  conc.reports[[i]]<- read.table(file.path(projFolder, "FASTQ", "CONCATENATED", conc.reports.files[i]))
}
```

```{r visualize concatenation reports}
par(oma=c(6,5,0,0))
labels <- gsub(".merged.fastq.gz.report","",conc.reports.files)
barplot(sapply(conc.reports,nrow), ylab="Concatenated files", names=labels, las=2, col=report.colours[1])
```

