# Case/Control Analysis
In case there is a casecontrol column in the file sampleInfo.txt we perform a search for causative mutations.

First, we identify possible candidate loci. For that, we create two sets of candidate loci, homozygeous reference and
homozygeous alternative. Hereby, we ignore the missing genotypes as well as accepting two wrong genotype across all
cases.

Then, we check if for those particular loci the controls do not have this particular homozygeous genotype (besides five
accepted mistakes).

```{r}
if(!is.null(sampleInfo)){
  if(sum(is.element(colnames(sampleInfo), "casecontrol"))>0){
    if(ncol(sampleInfo)>1){
      mmCase <- 2
      mmControl <- 5
      vcf.cases <- vcf$genotypes[sampleInfo$casecontrol=="case",]
      vcf.control <- vcf$genotypes[sampleInfo$casecontrol=="control",]
      
      vcf.cases.table <- apply(vcf.cases,2,table)
      vcf.control.table <- apply(vcf.control,2,table)
      
      nCases <- nrow(vcf.cases)
      nControl <- nrow(vcf.control)
      
      candidateLociRef <- c()
      candidateLociAlt <- c()
      for(i in 1:length(vcf.cases.table)){
        tmp <- vcf.cases.table[[i]]
        homRef <- tmp[names(tmp)=="00"] + tmp[names(tmp)=="03"]
        homAlt <- tmp[names(tmp)=="02"] + tmp[names(tmp)=="03"]
        if(length(homRef)==0) homRef <- 0
        if(length(homAlt)==0) homAlt <- 0
        if(homRef>=(nCases - mmCase)) candidateLociRef <- c(candidateLociRef, names(vcf.cases.table)[i])
        if(homAlt>=(nCases - mmCase)) candidateLociAlt <- c(candidateLociAlt, names(vcf.cases.table)[i])
      }
      
      causativeRef <- c()
      causativeAlt <- c()
      for(i in 1:length(candidateLociRef)){
        tmp <- vcf.control.table[[which(names(vcf.control.table)==candidateLociRef[i])]]
        homRef <- tmp[names(tmp)=="00"] + tmp[names(tmp)=="03"]
        homAlt <- tmp[names(tmp)=="02"] + tmp[names(tmp)=="03"]
        if(length(homRef)==0) homRef <- 0
        if(length(homAlt)==0) homAlt <- 0
      #  if(homRef>=(nControl - 1)) causativeRef <- c(causativeRef, names(vcf.control.table)[i])
        if(homAlt>=(nControl - mmControl)) causativeRef <- c(causativeRef, names(vcf.control.table)[i])
      }
      
      for(i in 1:length(candidateLociAlt)){
        tmp <- vcf.control.table[[which(names(vcf.control.table)==candidateLociAlt[i])]]
        homRef <- tmp[names(tmp)=="00"] + tmp[names(tmp)=="03"]
        homAlt <- tmp[names(tmp)=="02"] + tmp[names(tmp)=="03"]
        if(length(homRef)==0) homRef <- 0
        if(length(homAlt)==0) homAlt <- 0
        if(homRef>=(nControl - mmControl)) causativeAlt <- c(causativeAlt, names(vcf.control.table)[i])
      #  if(homAlt>=(nControl - 1)) causativeAlt <- c(causativeAlt, names(vcf.control.table)[i])
      }
    }
  }
}
```

Then we identify the following potentially causative mutations:

```{r}
if(!is.null(sampleInfo)){
  if(sum(is.element(colnames(sampleInfo), "casecontrol"))>0){
    if(ncol(sampleInfo)>1){
      if(length(causativeRef)>0){
        datatable(as.data.frame(causativeRef))
      } else {
        cat("No potentially mutations found for homozygeous reference!\n")
      }
      if(length(causativeAlt)>0){
        datatable(as.data.frame(causativeAlt))
      } else {
        cat("No potentially mutations found for homozygeous alternative!\n")
      }
    }
  }
}
```

