
## Reference Genome

### Basic stats

#### Filtered VCF
Basic stats and combinations for the called variants for filtered calls.

```{r}

if(refAvail) out <- getBasicStats(vcfRef)
if(refAvail) kable_styling(out[[1]], "striped", position = "left")  
if(refAvail) kable_styling(out[[2]], "striped", position = "left") 

```

#### Unfiltered VCF
Basic stats and combinations for the called variants for unfiltered calls.

```{r}
if(refAvail) out <- getBasicStats(vcfRefUnfiltered)
if(refAvail) kable_styling(out[[1]], "striped", position = "left")  
if(refAvail) kable_styling(out[[2]], "striped", position = "left") 
```

### Info Stats

#### Filtered VCF
Genotype stats for called variants for filtered calls.

```{r}
kable_styling(getInfoSummary(vcfRef), "striped", position = "left")  
```

#### Unfiltered VCF
Genotype stats for called variants for unfiltered calls.

```{r}
kable_styling(getInfoSummary(vcfRefUnfiltered), "striped", position = "left")  
```

### Visuals

#### SNP position vs snp index
Plot the snp position

```{r, fig.width=10}
par(mfrow=c(1,2))
plot(as.vector(as.matrix(vcfRef$map[,4])), ylab="Pos. on ref", xlab="SNP index", main="Filtered")
plot(as.vector(as.matrix(vcfRefUnfiltered$map[,4])), ylab="Pos. on ref", xlab="SNP index", main="Unfiltered")
```

#### SNP loci on reference
Distribution of the variant across the cluster

We line up here the chromosomes in the order as they appear in the created fasta index

```{r, fig.width=10}
par(mfrow=c(1,2))

positions <- as.data.frame(vcfRef$map)
for(i in 1:nrow(positions)){
  positions[i,4] <- positions[i,4] + refInfo[which(refInfo[,1]==positions[i,1]),3]
}
hist(as.vector(positions[,4]), main="SNP loci distribution (filtered)", xlab="Position", col=report.colours[1])

positions <- as.data.frame(vcfRefUnfiltered$map)

for(i in 1:nrow(positions)){
  positions[i,4] <- positions[i,4] + refInfo[which(refInfo[,1]==positions[i,1]),3]
}
hist(as.vector(positions[,4]), main="SNP loci distribution (unfiltered)", xlab="Position", col=report.colours[1])
```

#### Missing variants per samples (absolute)

Missing variant calls per sample

```{r, fig.width=10}
par(mfrow=c(1,2))
missingnessPerSample <- apply(vcfRef$genotypesInfo==".,.",2,sum)
barplot(missingnessPerSample, las=2, main="Missing variants per sample (filtered)", col=report.colours[1])

missingnessPerSample <- apply(vcfRefUnfiltered$genotypesInfo==".,.",2,sum)
barplot(missingnessPerSample, las=2, main="Missing variants per sample (unfiltered)", col=report.colours[1])
```

And that in relation to the sequencing depth

```{r, fig.width=10}
par(mfrow=c(1,2))
missingnessPerSample <- apply(vcfRef$genotypesInfo==".,.",2,sum)
plot(totalTrimmedSeq, missingnessPerSample, col=report.colours[1], xlab="Total sequences", ylab="Missing variants")

missingnessPerSample <- apply(vcfRefUnfiltered$genotypesInfo==".,.",2,sum)
plot(totalTrimmedSeq, missingnessPerSample, col=report.colours[1], xlab="Total sequences", ylab="Missing variants")
```

#### Missing variants per samples (percent)

```{r, fig.width=10}
par(mfrow=c(1,2))
barplot(apply(vcfRef$genotypesInfo==".,.",2,sum)/nrow(vcfRef$map)*100, las=2, main="Missing variants per sample (in %), filtered", col=report.colours[1], ylim=c(0,100))
barplot(apply(vcfRefUnfiltered$genotypesInfo==".,.",2,sum)/nrow(vcfRefUnfiltered$map)*100, las=2, main="Missing variants per sample (in %), unfiltered", col=report.colours[1], ylim=c(0,100))
```

And that in relation to the sequencing depth

```{r, fig.width=10}
par(mfrow=c(1,2))
missingnessPerSample <- apply(vcfRef$genotypesInfo==".,.",2,sum)/nrow(vcfRef$map)*100
plot(totalTrimmedSeq, missingnessPerSample, col=report.colours[1], xlab="Total sequences", ylab="Missing variants")

missingnessPerSample <- apply(vcfRefUnfiltered$genotypesInfo==".,.",2,sum)/nrow(vcfRefUnfiltered$map)*100
plot(totalTrimmedSeq, missingnessPerSample, col=report.colours[1], xlab="Total sequences", ylab="Missing variants")
```


```{r}
out <- apply(vcfRef$genotypesInfo==".,.",2,sum)/nrow(vcfRef$map)*100
tmp <- cbind(names(out), out)
rownames(tmp) <- NULL
colnames(tmp) <- c("Sample", "Missingness (preliminary mock)")

if(nrow(tmp)>0){
  rownames(tmp) <- 1:nrow(tmp)
 # colnames(tmp) <- c("Sample", "Total sequences")
  #knitr::kable(tmp) %>% kable_styling
  datatable(tmp, extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
}
```
