
# Variants

```{r}
getBasicStats <- function(x){
  output <- list()
  out <- c()
  out[1] <- dim(x$map)[1]
  names(out)[1] <- "Total number of variants"

  out[2] <- sum(sapply(apply(x$genotypes,2,table),length)==1)
  names(out)[2] <- "Monomorphic sites"

  out_html <- knitr::kable(out, col.names = NULL, "html")
  
  output[[1]] <- out_html
  
  output[[2]] <-  knitr::kable(table(x$map[,c(5,6)]), "html")
  
  output
  
}

getInfoSummary <- function(x){
  y <- x$info
  
  out <- rbind(summary(y[, "AC"]),
               summary(y[, "AF"]),
               summary(y[, "DP"]),
               summary(y[, "AV"]),
               summary(y[, "NS"]))
  
  rownames(out) <- c("Allele count",
                     "Allele frequency",
                     "Total depth",
                     "Average depth",
                     "Samples with data")
 
  colnames(out) <- c("Minimum",
                     "1st Quartile",
                     "Median",
                     "Mean",
                     "3rd Quartile",
                     "Maximum")
   
  out_html <- knitr::kable(out, "html")
    
  out_html
}

if(refAvail){
  refInfo.file <- paste0(refGenome.file,".fai")
  refInfo <- read.table(refInfo.file)
}
```

Import the Result vcfs

```{r import vcf, echo=FALSE, message=FALSE, warning=FALSE}
vcf <- importVCF(file.path(projFolder, "MPILEUP", "mpileup_finalMock", "GSC.vcf"), getInfo = TRUE)
vcfUnfiltered <- importVCF(file.path(projFolder, "MPILEUP", "mpileup_finalMock", "GSC.unfiltered.vcf"), getInfo = TRUE)

vcfGSC <- importVCF(file.path(projFolder, "FASTQ", "TRIMMED", "GSC.vcf"), getInfo = TRUE)
vcfGSCUnfiltered <- importVCF(file.path(projFolder, "FASTQ", "TRIMMED", "GSC.unfiltered.vcf"), getInfo = TRUE)

if(refAvail){
  vcfRef <- importVCF(file.path(projFolder, "MPILEUP", "mpileup_reference", "GSC.vcf"), getInfo = TRUE)
  vcfRefUnfiltered <- importVCF(file.path(projFolder, "MPILEUP", "mpileup_reference", "GSC.unfiltered.vcf"), getInfo = TRUE)
}
```

## Final mock

### Basic stats

Basic stats and combinations for the called variants for filtered calls.

```{r}
out <- getBasicStats(vcf)
  kable_styling(out[[1]], "striped", position = "left")  
  
  kable_styling(out[[2]], "striped", position = "left")  
```

Basic stats and combinations for the called variants for unfiltered calls.

```{r}
out <- getBasicStats(vcfUnfiltered)
  kable_styling(out[[1]], "striped", position = "left")  
  
  kable_styling(out[[2]], "striped", position = "left")  
```

### Info Stats

Genotype stats for called variants for filtered calls.

```{r}
kable_styling(getInfoSummary(vcf), "striped", position = "left")  
```

Genotype stats for called variants for unfiltered calls.

```{r}
kable_styling(getInfoSummary(vcfUnfiltered), "striped", position = "left")  
```


### Visuals

#### SNP position vs snp index
Plot the snp position

```{r, fig.width=10}
par(mfrow=c(1,2))
plot(as.vector(as.matrix(vcf$map[,4])), ylab="Pos. on mock", xlab="SNP index", main="Filtered")
plot(as.vector(as.matrix(vcfUnfiltered$map[,4])), ylab="Pos. on mock", xlab="SNP index", main="Unfiltered")
```

#### SNP loci on cluster
Distribution of the variant across the cluster

```{r, fig.width=10}
par(mfrow=c(1,2))
hist(as.vector(as.matrix(vcf$map[,4])), main="SNP loci distribution (filtered)", xlab="Position", col=report.colours[1])
hist(as.vector(as.matrix(vcfUnfiltered$map[,4])), main="SNP loci distribution (unfiltered)", xlab="Position", col=report.colours[1])
```

#### Missing variants per samples (absolute)

Missing variant calls per sample

```{r, fig.width=10}
par(mfrow=c(1,2))
missingnessPerSample <- apply(vcf$genotypesInfo==".,.",2,sum)
barplot(missingnessPerSample, las=2, main="Missing variants per sample (filtered)", col=report.colours[1])
missingnessPerSample <- as.matrix(missingnessPerSample)

missingnessPerSample <- apply(vcfUnfiltered$genotypesInfo==".,.",2,sum)
barplot(missingnessPerSample, las=2, main="Missing variants per sample (unfiltered)", col=report.colours[1])
missingnessPerSample <- as.matrix(missingnessPerSample)
```

#### Missing variants per samples (percent)

```{r, fig.width=10}
par(mfrow=c(1,2))
barplot(apply(vcf$genotypesInfo==".,.",2,sum)/nrow(vcf$map)*100, las=2, main="Missing variants per sample (in %), filtered", col=report.colours[1])
barplot(apply(vcfUnfiltered$genotypesInfo==".,.",2,sum)/nrow(vcfUnfiltered$map)*100, las=2, main="Missing variants per sample (in %), unfiltered", col=report.colours[1])
```
## Preliminary mock

### Basic stats

Basic stats and combinations for the called variants for filtered calls.

```{r}
out <- getBasicStats(vcfGSC)
kable_styling(out[[1]], "striped", position = "left")  

kable_styling(out[[2]], "striped", position = "left")  
```

Basic stats and combinations for the called variants for unfiltered calls.

```{r}
out <- getBasicStats(vcfGSCUnfiltered)
kable_styling(out[[1]], "striped", position = "left")  

kable_styling(out[[2]], "striped", position = "left") 
```

### Info Stats

Genotype stats for called variants for filtered calls.

```{r}
kable_styling(getInfoSummary(vcfGSC), "striped", position = "left")  
```

Genotype stats for called variants for unfiltered calls.

```{r}
kable_styling(getInfoSummary(vcfGSCUnfiltered), "striped", position = "left")  
```



### Visuals

#### SNP position vs snp index
Plot the snp position

```{r, fig.width=10}
par(mfrow=c(1,2))
plot(as.vector(as.matrix(vcfGSC$map[,4])), ylab="Pos. on mock", xlab="SNP index", main="Filtered")
plot(as.vector(as.matrix(vcfGSCUnfiltered$map[,4])), ylab="Pos. on mock", xlab="SNP index", main="Unfiltered")
```

#### SNP loci on cluster
Distribution of the variant across the cluster

```{r, fig.width=10}
par(mfrow=c(1,2))
hist(as.vector(as.matrix(vcfGSC$map[,4])), main="SNP loci distribution (filtered)", xlab="Position", col=report.colours[1])
hist(as.vector(as.matrix(vcfGSCUnfiltered$map[,4])), main="SNP loci distribution (unfiltered)", xlab="Position", col=report.colours[1])
```

#### Missing variants per samples (absolute)

Missing variant calls per sample

```{r, fig.width=10}
par(mfrow=c(1,2))
missingnessPerSample <- apply(vcfGSC$genotypesInfo==".,.",2,sum)
barplot(missingnessPerSample, las=2, main="Missing variants per sample (filtered)", col=report.colours[1])
missingnessPerSample <- as.matrix(missingnessPerSample)

missingnessPerSample <- apply(vcfGSCUnfiltered$genotypesInfo==".,.",2,sum)
barplot(missingnessPerSample, las=2, main="Missing variants per sample (unfiltered)", col=report.colours[1])
missingnessPerSample <- as.matrix(missingnessPerSample)
```

#### Missing variants per samples (percent)

```{r, fig.width=10}
par(mfrow=c(1,2))
barplot(apply(vcfGSC$genotypesInfo==".,.",2,sum)/nrow(vcfGSC$map)*100, las=2, main="Missing variants per sample (in %), filtered", col=report.colours[1])
barplot(apply(vcfGSCUnfiltered$genotypesInfo==".,.",2,sum)/nrow(vcfGSCUnfiltered$map)*100, las=2, main="Missing variants per sample (in %), unfiltered", col=report.colours[1])
```


## Reference Genome

### Basic stats

Basic stats and combinations for the called variants for filtered calls.

```{r}
if(refAvail){
out <- getBasicStats(vcfRef)
  kable_styling(out[[1]], "striped", position = "left")  
  
  kable_styling(out[[2]], "striped", position = "left") 
}
```

Basic stats and combinations for the called variants for unfiltered calls.

```{r}
if(refAvail){
  out <- getBasicStats(vcfRefUnfiltered)
    kable_styling(out[[1]], "striped", position = "left")  

    kable_styling(out[[2]], "striped", position = "left") 
}
```

### Info Stats

Genotype stats for called variants for filtered calls.

```{r}
kable_styling(getInfoSummary(vcfRef), "striped", position = "left")  
```

Genotype stats for called variants for unfiltered calls.

```{r}
kable_styling(getInfoSummary(vcfRefUnfiltered), "striped", position = "left")  
```

### Visuals

#### SNP position vs snp index
Plot the snp position

```{r, fig.width=10}
par(mfrow=c(1,2))
plot(as.vector(as.matrix(vcfRef$map[,4])), ylab="Pos. on ref", xlab="SNP index", main="Filtered")
plot(as.vector(as.matrix(vcfRefUnfiltered$map[,4])), ylab="Pos. on ref", xlab="SNP index", main="Unfiltered")
```

#### SNP loci on reference
Distribution of the variant across the cluster

We line here the chromosomes up in the order as they appear in the created fasta index

```{r, fig.width=10}
par(mfrow=c(1,2))

positions <- as.data.frame(vcfRef$map)
for(i in 1:nrow(positions)){
  positions[i,4] <- positions[i,4] + refInfo[which(refInfo[,1]==positions[i,1]),3]
}
hist(as.vector(positions[,4]), main="SNP loci distribution (filtered)", xlab="Position", col=report.colours[1])

positions <- as.data.frame(vcfRefUnfiltered$map)

for(i in 1:nrow(positions)){
  positions[i,4] <- positions[i,4] + refInfo[which(refInfo[,1]==positions[i,1]),3]
}
hist(as.vector(positions[,4]), main="SNP loci distribution (unfiltered)", xlab="Position", col=report.colours[1])
```

#### Missing variants per samples (absolute)

Missing variant calls per sample

```{r, fig.width=10}
par(mfrow=c(1,2))
missingnessPerSample <- apply(vcfRef$genotypesInfo==".,.",2,sum)
barplot(missingnessPerSample, las=2, main="Missing variants per sample (filtered)", col=report.colours[1])
missingnessPerSample <- as.matrix(missingnessPerSample)

missingnessPerSample <- apply(vcfRefUnfiltered$genotypesInfo==".,.",2,sum)
barplot(missingnessPerSample, las=2, main="Missing variants per sample (unfiltered)", col=report.colours[1])
missingnessPerSample <- as.matrix(missingnessPerSample)
```

#### Missing variants per samples (percent)

```{r, fig.width=10}
par(mfrow=c(1,2))
barplot(apply(vcfRef$genotypesInfo==".,.",2,sum)/nrow(vcfRef$map)*100, las=2, main="Missing variants per sample (in %), filtered", col=report.colours[1])
barplot(apply(vcfRefUnfiltered$genotypesInfo==".,.",2,sum)/nrow(vcfRefUnfiltered$map)*100, las=2, main="Missing variants per sample (in %), unfiltered", col=report.colours[1])
```